---
title: "Visualization 2"
output: github_document
date: "2025-09-30"
---

```{r }
library(tidyverse)
library(patchwork)
library(p8105.datasets)
library(viridis)
```

```{r}
data("weather_df")
```

Let's make our basic scatterplot
##labs( adds labels to graph
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color = name), alpha =0.5) +
  labs(
    x= "minimum daily temp",
    y= "maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  )
```

##scales: can change scale by x or y axis limits zooms into only a section of the plot that you care about.Scale_clor_hue changes color scheme, but Jeff wants us to use the "virdis" package instead because it's "friendlier"?
```{r}
weather_df |>
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color = name), alpha =0.5) +
  labs(
    x= "minimum daily temp",
    y= "maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) +
  scale_y_continuous(
    trans= "sqrt",
    limits = c(10,30)
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

##save a big code chunk as a name called ggp_temperature so you dont have to tyoe it out. Could also use ggsave to save the scatterplot
```{r}
ggp_temperature =
  weather_df |>
  ggplot(aes(x=tmin, y=tmax)) +
  geom_point(aes(color = name), alpha =0.5) +
  labs(
    x= "minimum daily temp",
    y= "maximum daily temp",
    title = "Temperature scatterplot",
    caption = "Data from NOAA",
    color = "Location"
  ) +
  scale_x_continuous(
    breaks = c(-20, 0, 25),
    labels = c("-20C", "0", "25")
  ) +
  scale_y_continuous(
    trans= "sqrt",
    limits = c(10,30)
  ) +
  viridis::scale_color_viridis(
    discrete = TRUE
  )
```

##theme minimal does white background with grey lines
```{r}
ggp_temperature +
  theme_minimal()+
  theme(lgend.position="bottom")
```

##Adding data in geoms
```{r}
central_park_df =
  weather_df |>
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |>
  filter(name == "Molokai_HI")

ggplot(data = molokai_df, aes(x = date, y= tmax, color = name)) +
  geom_point()+
  geom_line(data = central_park_df)
```

##patchwork: Make three plots and combine using patchwork to squeeze multiple plots in the same panel to have a lot of info without taking up a lot of space
```{r}
ggp_tmax_tmin =
  weather_df |>
  ggplot(aes(x=tmin, y=tmax, color=name)) +
  geom_point(alpha = 0.5) +
  theme(legend.position = "none")

ggp_prec_density =
  weather_df |>
  filter(prcp >0) |>
  ggplot (aes(x =prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  theme(legend.position = "none")

ggp_temp_season =
  weather_df |>
  ggplot(aes(x=date, y=tmax, color = name)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se=FALSE) +
  theme(legend.position = "bottom")

(ggp_tmax_tmin + ggp_prec_density) / ggp_temp_season
```

##Data manipulation
Let's make temperature violin plots. Factor relevel assigns molokai, cp, and waterhole in the order you want
```{r}
weather_df |>
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |>
  
  ggplot(aes(x=name, y=tmax, fill=name))+
  geom_violin(alpha = 0.5)
```

```{r}
weather_df |>
  mutate(name = fct_reorder(name, tmax)) |>
  ggplot(aes(x=name, y=tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

```{r}
pulse_df =
  haven::read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl :bdi_score_12m,
    names_to = "visit",
    names_prefix = "bdi_score_",
    values_to = "bdi"
  ) |>
  mutate(visit = fct_inorder(visit))

pulse_df |>
  ggplot(aes(x=visit, y=bdi)) +
  geom_boxplot()
```

Make a plot for the FAS study
```{r}
pups_df =
  read_csv("data/FAS_pups.csv", na = c("NA", ".", ""), skip = 3) |>
  janitor::clean_names() |>
  mutate(
    sex= case_match(
      sex,
      1~"male",
      2~"female"
    )
  )

litters_df =
  read_csv("data/FAS_litters.csv", na=c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "tx_day"), sep=3)

fas_df =
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |>
  select(pd_ears:tx_day)|>
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome",
    names_prefix = "pd_",
    values_to = "pn_day") |>
  drop_na() |>
  ggplot(aes(x = dose, y=pn_day)) +
  geom_violin() +
  facet_grid(tx_day ~ outcome)
  
```



